---
title: "STA365_homework2_code"
author: "Ruike Xu"
date: "3/29/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 2

```{r}
# install.packages("rjags")
# install.packages("coda")

set.seed(1006562550)
library(coda)
library(R2jags)
library(lattice)
```

$$ Warren:\ \ \tau(1-\theta)+\theta(1-\lambda) $$
$$ Trump(\psi):\ \ \lambda\theta + (1-\tau)(1-\theta) $$

```{r}

# Model construction with likelihood function and prior
model.JAGS = function(){
    y ~ dbinom(psi, n)
    theta ~ dbeta(3, 7)
    tau ~ dbeta(6, 4)
    lambda ~ dbeta(7, 3)
    psi <- theta * lambda + (1 - theta) * (1 - tau)
  }
```

The model we are constructing is intended to predict which candidate is likely to win the 2024 US presidential election (Donald Trump or Elizabeth Warren), which is based on binomial likelihood and beta prior. $\theta$ is the probability that the voter claims to vote for trump (might lying), thus, $1 - \theta$ is the probability that the person claims to vote for Warren. $\lambda$ is the percentage of the population that is actually willing to vote for Trump while $\tau$ is the percentage of the population that is actually willing to vote for Warren. We can detect the 'Quasi Liberal' and 'Quasi Republican' as $1-\lambda$ and $1-\tau$. To be more specific, the percentage of the population that claimed they would vote for Trump/Warren but ended voting the opposite. 

Therefore, we can derive that the actual probability of a person voting for Trump as $\psi:\lambda\theta + (1-\tau)(1-\theta)$ and the actual probability of a person voting for Warren as $\tau(1-\theta)+\theta(1-\lambda)$

To define $\tau$ and $\lambda$, we would assume more Republican people are less willing to claim their feelings in Liberal states, so $\tau$ ~ Beta(6, 4) and $\lambda$ ~ Beta(7,3)

```{r}
# Simulating data
n = 100000
y = 30000
data.JAGS = list(y = y, n = n)
```

```{r}
# Randomly select the initial values
inits.JAGS = function(){
  return(list(theta=rbeta(1, 3, 7),tau=rbeta(1, 6, 4),lambda=rbeta(1, 7, 3)))
}
```

```{r}
# Select parameters that will be simulated with MCMC model
para.JAGS = c("theta", "tau", "lambda", "psi")
```

```{r}
# Fit the MCMC model, number of iterations = 90000, burn in = 10000
fit.JAGS = jags(data=data.JAGS,inits=inits.JAGS,
                parameters.to.save = para.JAGS,
                n.chains=1,
                n.iter=90000,
                n.burnin=10000,
                model.file=model.JAGS)
```
```{r}
# Print model fit summary
print(fit.JAGS)
```
```{r}
# Generate traceplots
traceplot(fit.JAGS,mfrow=c(2,3),ask=FALSE)
```

```{r}
# Plot MCMC object to show all parameter densities along with traceplots
fit.JAGS.mcmc = as.mcmc(fit.JAGS)
plot(fit.JAGS.mcmc,ask=FALSE)
```
```{r}
# Summary as a MCMC object 
summary(fit.JAGS.mcmc)
```

```{r}
# Traceplots as a MCMC object
xyplot(fit.JAGS.mcmc,layout=c(2,3))
```

```{r}
# Density plot as a MCMC object
densityplot(fit.JAGS.mcmc)
```

### Part(d)

After we observe the results from part(c), the traceplots for the parameters seem fairly stationary, however, $\lambda, \tau, \theta$ seems a bit less appropriate than $\psi$, but I think the variations are acceptable. When we identify the density plots, all parameters are in appropriate shapes as the priors we set up. Therefore, I think the model we constructed seems reasonable without any adjustments. 


### Part(e)

The estimated probability of Trump winning the 2024 US Election is 0.236, which is lower than 0.3. Since we set the prior that people that claimed to vote for Warren but vote for Trump are slightly higher than people who claimed to vote for Trump but vote for Warren, we would expect more Quasi Liberal to vote for Trump. However, the prior setting does not influence much due to the huge likelihood probability difference.   

## Question 3
### Part(a)

$$\begin{aligned}
       \theta_1 = 1 + 0.065X_1 
          = 1 + 0.065 * 95 = 7.175 \\
       H_1 \sim Poisson(7.175) \\
\end{aligned}$$

```{r}
# Probability of Poisson distribution for at least 2 infections
P_atleast2 <- 1 - ppois(1, 7.175)
P_atleast2
```
Therefore, the probability that Indonesia observes at least 2 infections in 
the given time period is 0.994

### Part(b)
$$\begin{aligned}
       \theta_2 = 1 + 0.065X_2 
          = 1 + 0.065 * 150 = 10.75 \\
       H_2 \sim Poisson(10.75) \\
\end{aligned}$$

```{r}
# Probability of Poisson distribution for at least 28 infections
P_atleast28 <- 1 - ppois(27, 10.75)
P_atleast28
```

Therefore, the probability that Singapore observes 28 or more infections in the 
given time period is approximately 0. 

# ```{r}
# EQcount_per = mvspec(EQcount, log="no")
# predom3 = EQcount_per$details[order(EQcount_per$details [,3], 
#                               decreasing = TRUE)[1:3],]
# predom3
# 
# predom3_CI_up = 2*predom3[1:3,3]/qchisq(0.025,2)
# predom3_CI_low = 2*predom3[1:3,3]/qchisq(0.975,2)
# 
# result <- data.frame(Series=c(rep("EQcount", 3)), Predominant.Freq=c(predom3[1:3, 1]), Period = c(predom3[1:3, 2]), Spectrum=c(predom3[1:3, 3]), Lower = predom3_CI_low, Upper = predom3_CI_up)
# 
# result
# ```


```{r, echo=FALSE, fig.height = 3, fig.width = 5, fig.cap = "Seasonal time series plot for differencing of global annual earthquakes from 1900 to 2006"}
monthplot(diff(EQcount), main = "Seasonal plot after differencing", xlab="Time", ylab="Number of earthquakes")
```



```{r}
# Fit the logistic regression model with gambia data
gambia_predictor <- as.matrix(gambia%>%select(-8)) 

fit_JAGS_gambia_model = jags(
  data = gambia_predictor,
  inits = list(list(beta0 = 0,beta=c(rep(0, 7)),
  parameters.to.save = c("phc"),
  n.chains = 1,
  n.iter = 1000,
  n.burnin = 200,
  model.file = JAGS_spam_model
)
```

```{r}
# Fit the logistic regression model with gambia data
gambia_predictor <- as.matrix(gambia%>%select(-8)) 

fit_JAGS_gambia_model = jags(data = list(X = gambia_predictor, Y = gambia$y, 
                                         n = 2035, p = 7),
  inits = list(list(beta=c(rep(0, 7)), mu_a = 0, sigma_a2 = 0, mu_b = 0, 
                    sigma_b2 = 0)),
  parameters.to.save = c("phc"),
  n.chains = 1,
  n.iter = 1000,
  n.burnin = 200,
  model.file = JAGS_gambia_model
)
```